|

### 2023-05-16 paper/01-introduction.Rmd
some fixes

In recent years, tracking data have replaced traditional box-score statistics and play-by-play data as the state-of-the-art in sports analytics.  Numerous sports are collecting and releasing data on player and ball locations on the playing surface over the course of a game. This spatio-temporal multi-resolution source of data has provided exceptional opportunities for researchers to perform advanced studies at a more granular level to deepen our understanding of different sports.  For complete surveys on how tracking data have transformed sports analytics, see @Macdonald2020Recreating, @Baumer2023Big, and @Kovalchik2023Player.

@@

In an attempt to further develop analytics and innovate the game, the National Football League (NFL) introduced their player tracking system known as Next Gen Stats (NGS) in 2016 [@nfl2023ngs].
NGS uses radio frequency identification (RFID) chips placed in players' shoulder pads (and in the ball) to collect data at a rate of 10 frames per second.
The data captures real-time on-field information such as locations, speeds, and accelerations of all 22 players (as well as the ball).
While these data were initially only available for teams, media, and vendors, in December 2018 the NFL launched the inaugural edition of their annual Big Data Bowl (BDB) competition [@nfl2023big].

@@

\begin{table}
\caption{A summary of previously-existing pass rush metrics. \label{tab:metrics}}
\centering
\begin{tabular}{p{0.27\textwidth}p{0.67\textwidth}}
\hline
Metric & Description \\ 
\hline
Sacks & A defender tackles the QB behind the line of scrimmage before the QB is able to throw a pass \\
Hits & A defender tackles the QB behind the line of scrimmage after the QB is able to throw a pass. \\
Hurries & A defender pressures the QB behind the line of scrimmage forcing the QB to throw the ball prior sooner than intended. \\
Pressures & $\text{Hurries} + \text{ Hits } + \text{ Sacks}$ \\
Pass-Rush Productivity & $\displaystyle \frac{(\text{Hurries} + \text{Hits})/2 + \text{Sacks}}{\text{ Pass Rush Snaps}}$\\
Time In Pocket & Time (in seconds) between ball snap and throw or pocket collapse for a QB \\
NGS Get Off & Average time (in seconds) required for a defender to cross the line of scrimmage after the ball snap.\\
Pass Rush Win Rate & Rate at which pass rusher beats pass block within 2.5 seconds after ball snap\\
\hline
\end{tabular}
\end{table}

@@

In addition to these individual binary metrics, the sum of sacks, hits, and hurries is often defined as pressures.
This is better than the individual counts to some extent, but suffers from problems of subjectivity (e.g., whether there is an actual hurry or not).
Pro football focus (PFF) defines a metric called pass-rush productivity (PRP), which is a minor modification from the aforementioned pressures metric (see Table \ref{tab:metrics}).
In particular, PRP gives twice as much weight to a sack relative to hurries and hits, which is a small upgrade to pressures, however, the choice of weights is ad-hoc and still only copnsiders binary outcomes, similar to the shortcomings of previous metrics.

@@

More recently proposed metrics such as time in pocket (TIP), NGS get off [@Hermsmeyer2021nfl], and pass rush win rate (PRWR) [@Burke2018Created] are substantial improvements over the less sophisticated counting statistics, but nevertheless are still imperfect.
TIP refers to how long a QB can operate within the protected space behind the offensive line, known as the pocket.
However, this measure is highly context-dependent, as it can be influenced by a number of factors such as the defensive scheme or type of passing route.
NGS get off is an aggregated statistic, illustrating how quickly a defender can get past the line of scrimmage after the snap on average.
PRWR is created using player tracking data, which is at a more granular level than previous measures.
It demonstrates whether a pass rusher is able to beat their blocking matchup before a fixed time from the snap (2.5 seconds as chosen by ESPN).
However, this depends on the rather arbitrary time threshold used to define a pass rush win.  In addition, once a cutoff is chosen, PRWR converts continuous data to a win-loss indicator, becoming dichotomous like most of the metrics discussed above.

@@

In this paper, we focus on the evaluation of defensive linemen in football.
We propose STRAIN, a metric for measuring pass rush effectiveness, inspired by the concept of strain rate in materials science [@callister2009materials].
Our metric gives a continuous measure of pressure for every pass rusher on the football field over the course of an entire play.
This allows for the assessment of pass rush success even on plays that do not result in an observed outcome like a sack, hit, or hurry.
We view this as a major step forward for accurately evaluating defensive linemen performance.
Further, we note that although our focus in this paper is on defensive linemen, our proposed approach can be extended to the evaluation of offensive linemen in American football.

The remainder of this manuscript is outlined as follows.
We first describe the player tracking data provided by 2023 Big Data Bowl in Section \ref{sec:data}.
We then introduce the mathematical motivation and definition of our statistic STRAIN and our modeling approach in Section \ref{sec:methods}.
Next, we present applications of STRAIN and study different statistical properties of the metric in Section \ref{sec:results}.
We close with our discussion of future directions related to this work in Section \ref{sec:discussion}.

>>>>

In recent years, tracking data have replaced traditional box-score statistics and play-by-play data as the state of the art in sports analytics.
Numerous sports are collecting and releasing data on player and ball locations on the playing surface over the course of a game.
This spatio-temporal multi-resolution source of data has provided exceptional opportunities for researchers to perform advanced studies at a more granular level to deepen our understanding of different sports.
For complete surveys on how tracking data have transformed sports analytics, see @Macdonald2020Recreating, @Baumer2023Big, and @Kovalchik2023Player.

@@

In an attempt to foster analytics and innovate the game, the National Football League (NFL) introduced their player tracking system known as Next Gen Stats (NGS) in 2016 [@nfl2023ngs].
NGS uses radio frequency identification (RFID) chips placed in players' shoulder pads (and in the ball) to collect data at a rate of 10 frames per second.
The data captures real-time on-field information such as locations, speeds, and accelerations of all 22 players (as well as the ball).
While these data were initially only available for teams, media, and vendors, in December 2018 the NFL launched the inaugural edition of their annual Big Data Bowl (BDB) competition [@nfl2023big].

@@

\begin{table}
\caption{A summary of previously-existing pass rush metrics. \label{tab:metrics}}
\centering
\begin{tabular}{p{0.27\textwidth}p{0.67\textwidth}}
\hline
Metric & Description \\ 
\hline
Sacks & A defender tackles the QB behind the line of scrimmage before a QB throw \\
Hits & A defender tackles the QB behind the line of scrimmage after a QB throw \\
Hurries & A defender pressures the QB behind the line of scrimmage forcing the QB to throw the ball sooner than intended \\
Pressures & $\text{Hurries} + \text{ Hits } + \text{ Sacks}$ \\
Pass-Rush Productivity & $\displaystyle \frac{(\text{Hurries} + \text{Hits})/2 + \text{Sacks}}{\text{ Pass Rush Snaps}}$\\
Time In Pocket & Time (in seconds) between ball snap and throw or pocket collapse for a QB \\
NGS Get Off & Average time (in seconds) required for a defender to cross the line of scrimmage after the ball snap\\
Pass Rush Win Rate & Rate at which pass rusher beats pass block within 2.5 seconds after ball snap\\
\hline
\end{tabular}
\end{table}

@@

In addition, the sum of sacks, hits, and hurries is often defined as pressures.
This is better than the individual counts to some extent, but suffers from problems of subjectivity (e.g., whether there is an actual hurry or not).
Pro football focus (PFF) defines a metric called pass-rush productivity (PRP), which is a minor modification from the aforementioned pressures metric (see Table \ref{tab:metrics}).
In particular, PRP gives twice as much weight to a sack relative to hurries and hits, which is a small upgrade to pressures.
However, the choice of weights is ad-hoc and still only considers binary outcomes, similar to the shortcomings of previous metrics.

@@

More recently proposed metrics such as time in pocket (TIP), NGS get off [@Hermsmeyer2021nfl], and pass rush win rate (PRWR) [@Burke2018Created] are substantial improvements over the less sophisticated counting statistics, but nevertheless are still imperfect.
TIP refers to how long a QB can operate within the protected space behind the offensive line, known as the pocket.
However, this measure is highly context-dependent, as it can be influenced by a number of factors such as the defensive scheme or type of passing route.
NGS get off is an aggregated statistic, illustrating how quickly a defender can get past the line of scrimmage after the snap on average.
PRWR is created using player tracking data, which is at a more granular level than previous measures.
It demonstrates whether a pass rusher is able to beat their blocking matchup before a fixed time from the snap (2.5 seconds as chosen by ESPN).
However, this depends on the rather arbitrary time threshold used to define a pass rush win.
Besides, once a cutoff is chosen, PRWR converts continuous data to a win-loss indicator, becoming dichotomous like most of the metrics discussed above.

@@

In this paper, we focus on the evaluation of defensive linemen in football.
We propose STRAIN, a metric for measuring pass rush effectiveness, inspired by the concept of strain rate in materials science.
Our metric gives a continuous measure of pressure for every pass rusher on the football field over the course of an entire play.
This allows for the assessment of pass rush success even on plays that do not result in an observed outcome like a sack, hit, or hurry.
We view this as a major step forward for accurately evaluating defensive linemen performance.
Further, we note that although our focus in this paper is on defensive linemen, our proposed approach can be extended to the evaluation of offensive linemen in American football.

The remainder of this manuscript is outlined as follows.
We first describe the player tracking data provided by the Big Data Bowl 2023 in Section \ref{sec:data}.
We then introduce the mathematical motivation and definition of our statistic STRAIN and our modeling approach in Section \ref{sec:methods}.
Next, we present applications of STRAIN and study different statistical properties of the metric in Section \ref{sec:results}.
We close with our discussion of future directions related to this work in Section \ref{sec:discussion}.

### 2023-05-16 paper/02-data.Rmd
some fixes

In the forthcoming analysis, we rely on the data from the 2023 NFL Big Data Bowl provided by the NFL Next Gen Stats tracking system.
The data corresponds to 8,557 passing plays across 122 games in the first eight weeks of the 2021 NFL regular season. 
For each play, we have information on the on-field location, speed, angle, direction, and orientation of each player on the field and the football at a rate of 10 Hz (i.e. ten measurements per second), along with event annotations for each frame such as ball snap, pass forward, and quarterback sack, to name a few.

@@

Along with the tracking information, the 2023 BDB includes scouting data provided by Pro Football Focus (PFF). 
This contains manually-collected player-level information, such as the player's role (e.g., whether they are a pass rusher and pass blocker) and credited events (e.g., player is credited with hitting the QB on the play).
In this manuscript, we use the PFF data to identify 36,362 unique pass rushing attempts by players designated in the "pass rush" role across all plays.  <!-- Need to reword this.  There aren't 36,362 unique players.  It's like players by rushes or something like that.  -->
For context, Figure \ref{fig:pass_rushers_count} displays the distribution of the number of observed pass rushers involved in a play, ranging from 1 to 8 with 4 pass rushers (i.e. 4-man rush formation) as the most common value.
We also leverage the PFF scouting data to count how many hits, hurries, and sacks each pass rusher is credited with across the span of observed data. 
Additionally, we use the PFF player roles to identify the blocking matchup for each pass rusher in order to adjust for opponent strength, as discussed in Section \ref{sec:methods}.

>>>>

In the forthcoming analysis, we rely on the data from the NFL Big Data Bowl 2023 provided by the NFL Next Gen Stats tracking system.
The data corresponds to 8,557 passing plays across 122 games in the first eight weeks of the 2021 NFL regular season. 
For each play, we have information on the on-field location, speed, angle, direction, and orientation of each player on the field and the football at a rate of 10 Hz (i.e. 10 measurements per second), along with event annotations for each frame such as ball snap, pass forward, and quarterback sack, to name a few.

@@

Along with the tracking information, the 2023 BDB includes scouting data provided by Pro Football Focus (PFF). 
This contains manually-collected player-level information, such as the player's role (e.g., whether they are a pass rusher and pass blocker) and credited events (e.g., player is credited with hitting the QB on the play).
In this manuscript, we use the PFF data to identify 36,362 unique pass rush attempts by players designated in the "pass rush" role across all plays.
For context, Figure \ref{fig:pass_rushers_count} displays the distribution of the number of observed pass rushers involved in a play, ranging from 1 to 8 with 4 pass rushers (i.e. 4-man rush formation) as the most common value.
We also leverage the PFF scouting data to count how many hits, hurries, and sacks each pass rusher is credited with across the span of observed data. 
Additionally, we use the PFF player roles to identify the blocking matchup for each pass rusher in order to adjust for opponent strength, as discussed in Section \ref{sec:methods}.

### 2023-05-16 paper/03-methods.Rmd
some fixes

In materials science, strain is the deformation of a material from stress, showing the rate of change in a material's length relative to its original length [@strain].
Formally, let $L(t)$ be the distance between any given two points of interest within a material at time $t$, and $L_0$ be the initial distance between those two points.
The **strain** for a material at time $t$ is defined as 
\begin{equation}
s(t)= \frac{L(t)-L_{0}}{L_{0}}.
\end{equation}
Notice that this measure is unitless due to being a ratio of two quantities having the same dimension.

Accordingly, the **strain rate** of a material measures the change in its deformation with respect to time.
Mathematically, the strain rate of a material can be expressed as the derivative of its strain.
That is,
\begin{equation}
s'(t) = \frac{ds}{dt} = \frac{v(t)}{L_{0}},
\end{equation}
where $v(t)$ is the velocity at which the two points of interest within the material are moving away from or towards each other.
Whereas strain has no units, the strain rate is measured in inverse of time, usually 1/second.

@@

We consider a normal distribution to obtain the random intercepts for each player toward their respective position group means.
This provides us with average individual effects on STRAIN for each player while applying shrinkage, which is a useful property in this problem's context since we do not observe the same number of plays for each player.
With regards to the fixed-effects about pass rusher $j$ in play $i$, we include the length of the play (i.e., the elapsed time between the ball snap and a QB event), the positions of pass rusher and nearest blocker, and the number of pass blockers assigned to the pass rusher based on the PFF data.
Table \ref{tab:pos} shows our positional categorization for the pass rush and pass block roles.
We implement the model using penalized likelihood via the `lme4` package in `R` [@bateslme4; @R2023Language]. Additionally, we generate estimates of uncertainty for the player-level random intercepts based on 95% confidence intervals generated by sampling from the posterior distribution <!-- Posterior distribution of what? --> via parametric bootstrapping with the `merTools` package [@knowlesmerTools; @gelman2006data].

>>>>

In materials science, strain [@callister2018materials] is the deformation of a material from stress, showing the rate of change in a material's length relative to its original length.
Formally, let $L(t)$ be the distance between any given two points of interest within a material at time $t$, and $L_0$ be the initial distance between those two points.
The strain for a material at time $t$ is defined as 
\begin{equation}
s(t)= \frac{L(t)-L_{0}}{L_{0}}.
\end{equation}
Notice that this measure is unitless due to being a ratio of two quantities having the same dimension.

Accordingly, the strain rate [@callister2018materials] of a material measures the change in its deformation with respect to time.
Mathematically, the strain rate of a material can be expressed as the derivative of its strain.
That is,
\begin{equation}
s'(t) = \frac{ds}{dt} = \frac{v(t)}{L_{0}},
\end{equation}
where $v(t)$ is the velocity at which the two points of interest within the material are moving away from or towards each other.
Whereas strain has no units, the strain rate is measured in inverse of time, usually 1/second.

@@

We consider a normal distribution to obtain the random intercepts for each player toward their respective position group means.
This provides us with average individual effects on STRAIN for each player while applying shrinkage, which is a useful property in this problem's context since we do not observe the same number of plays for each player.
With regards to the fixed-effects about pass rusher $j$ in play $i$, we include the length of the play (i.e., the elapsed time between the ball snap and a QB event), the positions of pass rusher and nearest blocker, and the number of pass blockers assigned to the pass rusher based on the PFF data.
Table \ref{tab:pos} shows our positional categorization for the pass rush and pass block roles.
We implement the model using penalized likelihood via the `lme4` package in `R` [@bateslme4; @R2023Language]. Additionally, we generate estimates of uncertainty for the player-level random intercepts based on 95% confidence intervals generated by sampling from the posterior distribution <!-- Posterior distribution of what? --> via parametric bootstrapping of 10,000 samples with the `merTools` package [@knowlesmerTools; @gelman2006data].

### 2023-05-16 paper/04-results.Rmd
some fixes

Since STRAIN is observed continuously for every play in our data, this allows us to aggregate across all frames played and compute the average STRAIN for NFL pass rushers over the course of the eight-week sample size, as discussed in Section \ref{sec:defn}.
Based on the clearly distinct patterns for different positions as previously observed, we evaluate interior pass rushers (NT's and DT's) separately from edge rushers (OLB's and DE's).
Tables \ref{tab:edge} and \ref{tab:interior} are leaderboards for the NFL's best edge and interior rushers (with at least 100 plays) rated by the average STRAIN across all frames for the first eight weeks of the 2021 regular season.
The tables also consist of the total number of hits, hurries, and sacks for each defender.

<!-- Our results show many true positives with no surprises. -->
Our results are mostly consistent with conventional rankings of rushers.  For instance, Myles Garrett and TJ Watt are known as top-tier edge rushers and both show up in our top ten edge rushers. Additionally, Aaron Donald is nearly universally considered the best interior defender in football, and he appears at the top of our rankings of interior rushers.   
<!-- Rashan Gary, a young rusher with huge potential, ranks first among all pass rushers according to our metric over the period of time observed. -->
Moreover, our leader boards largely match the rankings of experts in the field.
For instance, there is considerable overlap between our lists and PFF's edge [@Monson2022nfl] and interior [@Linsey2022nfl] rusher rankings released after the 2021 season.
This ultimately lends credibility to our proposed metric STRAIN as a measure of pass rushing effectiveness.

@@

Next, we examine different statistical properties of our proposed metric STRAIN.
Our focus here is to understand the stability and predictability of STRAIN, and we pose the following question:
How much does our metric vary from week to week?
In other words, is previous performance as measured by our metric predictive of future performance based on our metric?
Below, we attempt the answer these questions, and the following results are for pass rushers with at least 100 snaps played during the first 8 weeks of the 2021 NFL regular season.

>>>>

Since STRAIN is observed continuously for every play in our data, this allows us to aggregate across all frames played and compute the average STRAIN for NFL pass rushers over the course of the eight-week sample size, as discussed in Section \ref{sec:defn}.
Based on the clearly distinct patterns for different positions as previously observed, we evaluate interior pass rushers (NT's and DT's) separately from edge rushers (OLB's and DE's).
Tables \ref{tab:edge} and \ref{tab:interior} are leaderboards for the NFL's best edge and interior rushers (with at least 100 plays) rated by the average STRAIN across all frames for the first eight weeks of the 2021 regular season.
The tables also consist of the total number of hits, hurries, and sacks (determined from PFF scouting data) for each defender.

<!-- Our results show many true positives with no surprises. -->
<!-- Rashan Gary, a young rusher with huge potential, ranks first among all pass rushers according to our metric over the period of time observed. -->

Our results are mostly consistent with conventional rankings of rushers.
Notably, Myles Garrett and TJ Watt are widely recognized as top-tier edge rushers and both show up in our top edge rusher list; whereas Aaron Donald, who is undoubtedly the best interior defender in football, appears at the top of our interior rusher rankings.
Moreover, our leaderboards largely match the rankings of experts in the field.
For instance, there is considerable overlap between our lists and PFF's edge [@Monson2022nfl] and interior [@Linsey2022nfl] rusher rankings released after the 2021 season.
This ultimately lends credibility to our proposed metric STRAIN as a measure of pass rushing effectiveness.

@@

Next, we examine different statistical properties of our proposed metric STRAIN.
Our focus here is to understand the stability and predictability of STRAIN, and we pose the following question:
How much does our metric vary from week to week?
In other words, is previous performance predictive of future performance based on our metric?
Below, we attempt the answer these questions, and the following results are for pass rushers with at least 100 snaps played during the first 8 weeks of the 2021 NFL regular season.
